\documentclass[KSmain.tex]{subfiles} 
\begin{document} 
%--------------------------------------------------------------------------------------------%
\section{Creating Dates and Times}

Dates are created using date by providing integer values for year, month and day and times are created
using time using hours, minutes, seconds and microseconds.

Date and time manipulation is provided by a built-in Python module datetime. This chapter assumes that
datetime has been imported using import datetime as dt.

\begin{framed}
\begin{verbatim}
>>> import datetime as dt

>>> yr, mo, dd = 2012, 12, 21

>>> dt.date(yr, mo, dd)
datetime.date(2012, 12, 21)

>>> hr, mm, ss, ms= 12, 21, 12, 21

>>> dt.time(hr, mm, ss, ms)
dt.time(12,21,12,21)

\end{verbatim}
\end{framed}

Dates created using date do not allow times, and dates which require a time stamp can be created using
datetime, which combine the inputs from date and time, in the same order.
\begin{framed}
\begin{verbatim}
>>> dt.datetime(yr, mo, dd, hr, mm, ss, ms)
datetime.datetime(2012, 12, 21, 12, 21, 12, 21)

\end{verbatim}
\end{framed}
%-----------------------------------------------------------------%


\begin{verbatim}
>>> datetime64(’2013’)
numpy.datetime64(’2013’)

>>> datetime64(’201309’)
numpy.datetime64(’201309’)

>>> datetime64(’20130901’)
numpy.datetime64(’20130901’)

>>> datetime64(’20130901T12:
00’) # Time
numpy.datetime64(’20130901T12:
00+0100’)

>>> datetime64(’20130901T12:
00:01’) # Seconds
numpy.datetime64(’20130901T12:
00:01+0100’)

>>> datetime64(’20130901T12:
00:01.123456789’) # Nanoseconds
numpy.datetime64(’20130901T12:
00:01.123456789+0100’)
\end{verbatim}

Date or time units can be explicitly included as the second input. The final example shows that rounding
can occur if the date input is not exactly representable using the date unit chosen.
%-----------------------------------------------------------------------%
\begin{framed}
\begin{verbatim}
>>> datetime64(’20130101T00’,’
h’)
numpy.datetime64(’20130101T00:
00+0000’,’h’)
>>> datetime64(’20130101T00’,’
s’)
numpy.datetime64(’20130101T00:
00:00+0000’)
>>> datetime64(’20130101T00’,’
ms’)
numpy.datetime64(’20130101T00:
00:00.000+0000’)
>>> datetime64(’20130101’,’
W’)
numpy.datetime64(’20121227’)
\end{verbatim}
\end{framed}

NumPy datetimes can also be initialized from arrays.
%-----------------------------------------------------------------------%
\begin{framed}
\begin{verbatim}
>>> dates = array([’20130901’,’
20130902’],
dtype=’datetime64’)
>>> dates
array([’20130901’,
’20130902’],
dtype=’datetime64[D]’)
>>> dates[0]
numpy.datetime64(’20130901’)
\end{verbatim}
\end{framed}

MORE
\end{document}

%-----------------------------------------------------------------------%
\newpage
% PAGE 142-143
\subsection{14.2 Dates Mathematics}
Date-times and dates (but not times, and only within the same type) can be subtracted to produce a
timedelta, which consists of three values, days, seconds and microseconds. Time deltas can also be added
to dates and times compute different dates – although date types will ignore any information in the time
delta hour or millisecond fields.
>>> d1 = dt.datetime(yr, mo, dd, hr, mm, ss, ms)
>>> d2 = dt.datetime(yr + 1, mo, dd, hr, mm, ss, ms)
>>> d2d1
datetime.timedelta(365)


Date Unit Common Name Range Time Unit Common Name Range
Y Year 9.2  1018 years h Hour 1.0  1015 years
M Month 7.6  1017 years m Minute 1.7  1013 years
W Week 1.7  1017 years s Second 2.9  1012 years
D Day 2.5  1016 years ms Millisecond 2.9  109 years

us Microsecond 2.9  106 years
ns Nanosecond 292 years
ps Picosecond 106 days
fs Femtosecond 2.6 hours
as Attosecond 9.2 seconds
Table 14.1: NumPy datetime64 range. The absolute range is January 1, 1970 plus the range.

>>> d2 + dt.timedelta(30,0,0)
datetime.datetime(2014, 1, 20, 12, 21, 12, 20)
>>> dt.date(2012,12,21) + dt.timedelta(30,12,0)
datetime.date(2013, 1, 20)
If times stamps are important, date types can be promoted to datetime using combine and a time.
>>> d3 = dt.date(2012,12,21)
>>> dt.datetime.combine(d3, dt.time(0))
datetime.datetime(2012, 12, 21, 0, 0)

Values in dates, times and datetimes can be modified using replace through keyword arguments.

>>> d3 = dt.datetime(2012,12,21,12,21,12,21)
>>> d3.replace(month=11,day=10,hour=9,minute=8,second=7,microsecond=6)
datetime.datetime(2012, 11, 10, 9, 8, 7, 6)
%-----------------------------------------------------------------------%
\newpage
\subsection{14.3 Numpy datetime64}
Version 1.7.0 of NumPy introduces a NumPy native datetime type known as datetime64 (to distinguish it
from the usual datetime type). The NumPy datetime type is considered experimental and is not fully supported
in the scientific python stack at the time of writing these notes. This said, it is already widely used
and should see complete support in the near future. Additionally, the native NumPy data type is generally
better suited to data storage and analysis and extends the Python datetime with additional features such
as business day functionality.

NumPy contains both datetime (datetime64) and timedelta (timedelta64) objects. These differ from
the standard Python datetime since they always store the datetime or timedelta using a 64-bit integer plus
a date or time unit. The choice of the date/time unit affects both the resolution of the datetime as well
as the permissible range. The unit directly determines the resolution - using a date unit of a day (’D’)
limits to resolution to days. Using a date unit of a week (’W’) will allow a minimum of 1 week difference.
Similarly, using a time unit of a second (’s’) will allow resolution up to the second (but not millisecond).
The set of date and time units, and their range are presented in Table 14.1.

NumPy datetimes can be initialized using either human readable strings or using numeric values. The
string initialization is simple and datetimes can be initialized using year only, year and month, the complete
date or the complete date including a time (and optional timezone). The default time resolution is
nanoseconds ($10^-9$) and T is used to separate the time from the date.
\end{document}

