
scipy.stats.linregress(x, y=None)[source]
Calculate a regression line

This computes a least-squares regression for two sets of measurements.

Parameters:	
x, y : array_like
two sets of measurements. Both arrays should have the same length. If only x is given (and y=None), then it must be a two-dimensional array where one dimension has
scipy.stats.linregress(x, y=None)[source]
Calculate a regression line

This computes a least-squares regression for two sets of measurements.

Parameters:	
x, y : array_like
two sets of measurements. Both arrays should have the same length. If only x is given (and y=None), then it must be a two-dimensional array where one dimension has length 2. The two sets of measurements are then found by splitting the array along the length-2 dimension.
Returns:
\begin{itemize}	
\item slope : float
slope of the regression line
\item intercept : float
intercept of the regression line
\item r-value : float
correlation coefficient
\item \texttt{p-value} : float
two-sided p-value for a hypothesis test whose null hypothesis is that the slope is zero.
\item \texttt{stderr} : float
Standard error of the estimate
\end{itemize}
%----------------------------------------------------------------------------------------- %
\subsubsection*{Examples}

\begin{framed}
\begin{verbatim}
>>> from scipy import stats
>>> import numpy as np
>>> x = np.random.random(10)
>>> y = np.random.random(10)
>>> slope, intercept, r_value, p_value, std_err = stats.linregress(x,y)
\end{verbatim}
\end{framed}
To get coefficient of determination (\texttt{r\_squared})

\begin{framed}
\begin{verbatim}
>>>
>>> print "r-squared:", r_value**2
$ r-squared: 0.15286643777 sis test whose null hypothesis is that the slope is zero.
\item stderr : float
Standard error of the estimate

\subsubsection*{Examples}

>>>
>>> from scipy import stats
>>> import numpy as np
>>> x = np.random.random(10)
>>> y = np.random.random(10)
>>> slope, intercept, r_value, p_value, std_err = stats.linregress(x,y)
# To get coefficient of determination (r_squared)

>>> $
>>> print "r-squared:", r_value**2
r-squared: 0.15286643777